---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---
# Activity Monitoring

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r activity}
library(readr)
library(dplyr)
activity <- read_csv("~/Documents/Data Science/R/Coursera/activity.csv")
head(activity)
summary(activity)
plot(activity$date, activity$steps)
```

## Transforming the data/Histogram

```{r aggregate by date}
agg_steps <-  aggregate(steps ~ date, data=activity, FUN=sum)
colnames(agg_steps) <- c("Date","Steps")

hist(agg_steps$Steps , xlab="Number of steps", main="Total steps per day")
```

## What is mean total number of steps taken per day?

```{r}
(mean(agg_steps$Steps,na.rm=TRUE))
(median(agg_steps$Steps, na.rm=TRUE))
```
The average number of steps and median steps taken each day was `r as.integer(mean(agg_steps$Steps,na.rm=TRUE))` and `r as.integer(median(agg_steps$Steps,na.rm=TRUE))` steps respectively.


## What is the average daily activity pattern?
```{r}
library(ggplot2)
ggplot(agg_steps, aes(agg_steps$Date,agg_steps$Steps)) +
        geom_line(data=agg_steps, aes(y=agg_steps$Steps)) +
        xlab("date") + ylab("Avg. steps per day")
```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}

steps_interval<- aggregate(steps ~ interval, data=activity, FUN=sum)
steps_interval[which.max(steps_interval$steps),1]

plot(steps_interval$interval, steps_interval$steps, type='l', xlab='Interval', ylab='Number of steps', main="Average steps per interval")
```

## Imputing missing values
```{r}
sum(is.na(activity$steps)==TRUE)
```
Total missing values in dataset are `r sum(is.na(activity$steps)==TRUE)` . 

```{r}
activity_new<- activity
nas<- is.na(activity_new$steps)
avg_interval<- tapply(activity_new$steps, activity_new$interval, mean, na.rm=TRUE, simplify = TRUE)
activity_new$steps[nas] <- avg_interval[as.character(activity_new$interval[nas])]
names(activity_new)

sum(is.na(activity_new))
Total_Steps<- activity_new%>%
        group_by(date)%>%
        summarise(total_steps = sum(steps, na.rm=TRUE))
Total_Steps

#histogram after imputation
ggplot(Total_Steps, aes(x = total_steps)) +
        geom_histogram(fill = "blue", binwidth = 1000) +
        labs(title = "Daily Steps after imputation", x = "Interval", y = "No. of Steps")
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity_new$day <- factor(ifelse(weekdays(activity_new$date)=='Sunday' | weekdays(activity_new$date)=='Saturday', "Weekend", "Weekday"))
head(activity_new)
str(activity_new)
```

## Panel plot for weekdays and weekends
```{r}

Interval<- activity_new %>%
        group_by(interval, day) %>%
        summarise(avg_steps = mean(steps, na.rm=TRUE))
head(Interval)

plot<- ggplot(Interval, aes(x =interval , y=avg_steps, color=day)) +
       geom_line() +
       labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") +
       facet_wrap(~day, ncol = 1, nrow=2)
print(plot)
```

From the plots and the data , it is clear that people are active over the weekend compared to the weekdays which makes us understand that people are less active during work hours.














